---
title: "Institutions, Reforms and Economic Growth"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```

## Data cleaning and preparation

This was quite the learning experience. I chose to only work with the new EU member state data.

```{r}
library(readxl)
library(tidyverse)
library(stringr)
library(dplyr)
library(plm)
library(modelr)
library(corrplot)
library(stargazer)
library(gplots)
library(sjPlot)

url1 <- "https://github.com/klimaviu/Institutions-and-Growth/raw/master/Reforms_database_Project_2020.xlsx"
destfile1 <- "Reforms_database_Project_2020.xlsx"
curl::curl_download(url1, destfile1)

Reforms <- read_excel(destfile1,
                      skip = 1,
                      col_types = c(rep("text",2), rep("numeric", 81)))[2:716,] %>% 
  rename(New_old = `New/old`)

Reforms <- Reforms %>% mutate(
  EU =  factor(EU, levels = c("0", "1")),
  gov_party = factor(gov_party, levels = c("1", "2", "3", "4", "5"))
)

url2 <- "https://github.com/klimaviu/Institutions-and-Growth/raw/master/heritage-economic-freedom.xlsx"
destfile2 <- "heritage_economic_freedom.xlsx"
curl::curl_download(url2, destfile2)

heritage <- read_excel(destfile2,
                       col_types = c("text", rep("numeric", 14))) %>% 
  filter(Name %in% c(Reforms$Country)) %>% 
  rename(Year = "Index Year",
         Country = "Name",
         Overall_score = "Overall Score",
         Property_rights = "Property Rights",
         Judicial_effectiveness = "Judicial Effectiveness",
         Gov_integrity = "Government Integrity",
         Tax_burden = "Tax Burden",
         Gov_spending = "Government Spending",
         Fiscal_health = "Fiscal Health",
         Business_freedom = "Business Freedom",
         Labor_freedom = "Labor Freedom",
         Monetary_freedom = "Monetary Freedom",
         Trade_freedom = "Trade Freedom",
         Investment_freedom = "Investment Freedom",
         Financial_freedom = "Financial Freedom")

non_EU_countries <- c("United States", "Norway", "Switzerland", "Iceland", "Canada", "Japan", "Australia", "New Zealand", "Korea")         

'%!in%' <- function(x,y)!('%in%'(x,y))

Updated <- full_join(Reforms, heritage, by = c("Year", "Country")) %>% 
  filter(Year >=1996, Year <=2015) %>% 
  filter(Country %!in% non_EU_countries) %>% 
  select(New_old, Country, Year, GDP, GDP_per_capita, GDP_growth,
         trade_openness, capital_openness, Overall_score, Property_rights,
         Gov_integrity, Tax_burden, Gov_spending, Business_freedom, Labor_freedom,
         Monetary_freedom, Trade_freedom, Investment_freedom, Financial_freedom,
         lmr_fraser, GG_debt, EU, gov_party, gov_type, gov_new) %>% 
  mutate(
    Monetary_reform = c(rep(NA,3),ifelse((diff(Monetary_freedom,3))>0,1,0)), #New variables for reforms
    Spending_reform = c(rep(NA,3),ifelse((diff(Gov_spending,3))>0,1,0)),
    Tax_reform = c(rep(NA,3),ifelse((diff(Tax_burden,3))<0,1,0)),
    Financial_reform = c(rep(NA,3),ifelse((diff(Financial_freedom,3))>0,1,0)),
    Trade_reform = c(rep(NA,3),ifelse((diff(trade_openness,3))>0,1,0)),
    Capital_reform = c(rep(NA,3),ifelse((diff(capital_openness,3))>0,1,0)),
    Investment_reform = c(rep(NA,3),ifelse((diff(Investment_freedom,3))>0,1,0)),
    Overall_reform = c(rep(NA,3),ifelse((diff(Overall_score,3))>0,1,0)),
    lMon_ref = lag(Monetary_reform, 1), #Lagged reforms for the compution of correlations
    lSp_ref = lag(Spending_reform, 1),
    lTax_ref = lag(Tax_reform,1),
    lFin_ref = lag(Financial_reform, 1),
    lTr_ref = lag(Trade_reform,1),
    lCap_ref = lag(Capital_reform, 1),
    Crisis = factor(ifelse(Year %in% c(2008, 2009), 1, 0)) #Banking crisis dummy
  ) %>% 
  filter(Year >=1999, Year <=2015)

NMS <- Updated %>% filter(New_old == "NMS")
```

## Figure 1: Trade per capita and GDP per capita

```{r}
NMS %>% 
  mutate(
    population = GDP/GDP_per_capita,
    trade_per_capita = (trade_openness*GDP)/population
  ) %>% 
  ggplot()+
  geom_line(aes(Year, GDP_per_capita, color = Country))+
  geom_line(aes(Year, trade_per_capita))+
  facet_wrap(~Country, nrow = 2)+
  ylab("GDP per capita, Trade per capita")
```

## Figure 2: Heterogeneity accross countries

```{r}
plotmeans(GDP_growth ~ Country, main = "Heterogeneity across countries", data = NMS)
```

## Figure 3: Heterogeneity accross years

```{r}
plotmeans(GDP_growth ~ Year, main = "Heterogeneity across years", data = NMS)
```

## Figure 4: A correlations graph

```{r}
NMS_cor6 <- NMS %>%
  select(GDP_growth, Monetary_reform,Financial_reform, 
         Trade_reform, Gov_integrity, trade_openness)

corrplot(cor(NMS_cor6), type = "upper", order = "hclust")

final_correlations <- cor(NMS_cor6)
```

# *Fixed effects regressions*

## Looking at trade openness

```{r}
fixed <- plm(GDP_growth ~ trade_openness + Crisis,
             data=NMS,
             index=c("Country", "Year"),
             model = "within")
summary(fixed, se = vcovHC(fixed))

fixed2 <- plm(GDP_growth ~ trade_openness + factor(gov_party) + Gov_integrity + Crisis,
             data=NMS,
             index=c("Country", "Year"),
             model = "within")
summary(fixed2, se = vcovHC(fixed2))

```


## Looking at reforms:

```{r}
fixed_reform <- plm(GDP_growth ~ Trade_reform + Crisis,
                    data=NMS,
                    index=c("Country", "Year"), 
                    model = "within")
summary(fixed_reform, se = vcovHC(fixed_reform))
```

## Introducing more reform variables

```{r}
fixed_reform2 <- plm(GDP_growth ~ Trade_reform + Financial_reform + Monetary_reform + Crisis,
                     data=NMS,
                     index=c("Country", "Year"), 
                     model = "within")
summary(fixed_reform2, se = vcovHC(fixed_reform2))
```


## Introducing gov_type and Gov_integirty as a control variable

```{r}
fixed_reform3 <- plm(GDP_growth ~ Trade_reform + Financial_reform + Monetary_reform + Crisis + factor(gov_party) + Gov_integrity,
                     data=NMS,
                     index=c("Country", "Year"), 
                     model = "within")
summary(fixed_reform3, se = vcovHC(fixed_reform3))
```

## Lagging reforms:

```{r}
fixed_reform4 <- plm(GDP_growth ~ lag(Trade_reform) + lag(Financial_reform) + lag(Monetary_reform) + Crisis + factor(gov_party) + Gov_integrity,
                     data=NMS,
                     index=c("Country", "Year"), 
                     model = "within")
summary(fixed_reform4, se = vcovHC(fixed_reform4))
```

## Lagging one reform variable:

```{r}
fixed_reform5 <- plm(GDP_growth ~ Trade_reform + lag(Financial_reform) + Monetary_reform + Crisis + factor(gov_party) + Gov_integrity,
                     data=NMS,
                     index=c("Country", "Year"), 
                     model = "within")
summary(fixed_reform5, se = vcovHC(fixed_reform5))
```

## Generating heteroskedasticity robust SEs and the table

```{r}
rob_se1 <- list(sqrt(diag(vcovHC(fixed))),
               sqrt(diag(vcovHC(fixed_reform))),
               sqrt(diag(vcovHC(fixed_reform2))))

rob_se2 <- list(sqrt(diag(vcovHC(fixed_reform3))),
               sqrt(diag(vcovHC(fixed_reform4))),
               sqrt(diag(vcovHC(fixed_reform5))))

table1 <- stargazer(fixed, fixed_reform, fixed_reform2, 
                   digits = 3,
                   header = FALSE,
                   type = "text", 
                   se = rob_se1,
                   title = "Linear Panel Regression Models of the Impact of Institutions and Reforms on Economic Growth",
                   model.numbers = FALSE,
                   column.labels = c("(1)", "(2)", "(3)"))

table2 <- stargazer(fixed_reform3, fixed_reform4, fixed_reform5, 
                    digits = 3,
                    header = FALSE,
                    type = "text", 
                    se = rob_se2,
                    title = "Linear Panel Regression Models of the Impact of Institutions and Reforms on Economic Growth",
                    model.numbers = FALSE,
                    column.labels = c( "(4)", "(5)", "(6)"))
```